---
title: "p8105_hw1_yz5248"
author: "yz5248"
date: "2025-09-13"
output: github_document
---


```{r echo=TRUE}
library("moderndive")
data("early_january_weather")
```
```{r echo=TRUE}
nrow(early_january_weather)
ncol(early_january_weather)
round(mean(early_january_weather$temp), 1)
```

## Problem 1: Dataset Description

The `early_january_weather` dataset comes from the **moderndive** package and contains hourly weather data collected at LaGuardia Airport (LGA) in early January 2013.

This dataset has **`r nrow(early_january_weather)`** rows and **`r ncol(early_january_weather)`** columns.  
Key variables include:

- `time_hour`: date-time object indicating the hour of observation  
- `temp`: temperature in Fahrenheit  
- `humid`: relative humidity  
- `wind_speed`: wind speed in mph  
- `precip`: precipitation in inches  
- `visib`: visibility in miles  

The mean temperature during this period was **`r round(mean(early_january_weather$temp), 1)` °F**.

## Problem 1: Scatterplot of Temperature vs Time

```{r scatterplot, fig.width=7, fig.height=5}
library(ggplot2)

# Create scatterplot
scatter_temp_time <- ggplot(early_january_weather, aes(x = time_hour, y = temp, color = humid)) +
  geom_point() +
  labs(
    title = "Temperature vs Time (Colored by Humidity)",
    x = "Time (Hour)",
    y = "Temperature (°F)",
    color = "Humidity"
  ) +
  theme_minimal()

# Print plot so it shows up in the knitted document
scatter_temp_time

# Save plot as PNG file in your project directory
ggsave("scatter_temp_time.png", scatter_temp_time, width = 7, height = 5)
```

The scatterplot shows hourly temperature over time, with point colors indicating humidity levels. Several patterns are apparent:  
1. Daily cycles: Temperatures follow a roughly repeating daily pattern — lower in the early morning hours and higher during the day.  
2. Gradual warming: Overall temperatures tend to increase over the two-week period, with the warmest temperatures occurring around January 13–14.  
3. Humidity relationship: Points with darker colors (lower humidity) generally correspond to higher temperatures, while lighter blue points (higher humidity) appear at cooler temperatures.  
4. Variation: There is some day-to-day variability, but the general trend shows a gradual rise followed by a slight drop near the end.  


## Problem 2: Variable Types and Coercion

```{r create-df}
library(tidyverse)
set.seed(123) # for reproducibility

df <- tibble::tibble(
  norm_sample = rnorm(10),                      # numeric: random N(0,1)
  positive = rnorm(10) > 0,                     # logical: TRUE/FALSE
  char_vec    = sample(letters[1:3], 10, TRUE), # character vector
  factor_vec  = factor(sample(c("low","med","high"), 10, TRUE))
)

df
mean(df |> pull(norm_sample))  # works: numeric data
mean(df |> pull(positive))  # works: TRUE=1, FALSE=0 -> proportion > 0
```
```{r create numeric}
as.numeric(df$positive)
as.numeric(df$char_vec)
as.numeric(df$factor_vec)
```